---
title: "Midterm"
author: "Nishat"
format:
  html:
    embed-resources: true
execute:
  cache: true
---

## Setup

```{r}
#| message: false
#install.packages("tidyverse")
library(tidyverse)

library(glue)

#install.packages("rio")
library(rio)

library(fs)

#install.packages("patchwork")
library(patchwork)
```

```{r}
download_file <- function(year) {
  link <- glue(
    "https://www.oregon.gov/ode/educator-resources/assessment/TestResults20{year}/pagr_schools_ela_raceethnicity_{year-1}{year}.xlsx"
  )
  
  rio::import(
    link,
    setclass = "tibble",
    na = c("-", "--", "*"))}
```

# Part A.1 Downloading Clean Data

```{r}
raw_data <- map_dfr(15:18, download_file)

clean_data <- raw_data |>
  filter(`Student Group` %in% c("White", "Hispanic/Latino")) |>
  select(
    academic_year = `Academic Year`,
    district = District,
    school = School,
    student_group = `Student Group`,
    grade_level = `Grade Level`,
    `Number Level 1`,
    `Number Level 2`,
    `Number Level 3`,
    `Number Level 4`
  ) |>
  pivot_longer(
    cols = starts_with("Number Level"),
    names_to = "level",
    values_to = "n"
  ) |>
  mutate(
    level = str_remove(level, "Number Level "),
    n = as.integer(n)
  ) |>
  drop_na(n)

clean_data
```

# Part A.2 Summarizing by district, year, group, and level

```{r}
new_data <- clean_data |>
  group_by(academic_year, district, student_group, level) |>
  summarise(
    n = sum(n),
    n_schools = n_distinct(school),
    .groups = "drop"
  ) |>
  mutate(
    student_group = str_to_lower(student_group),
    student_group = str_replace_all(student_group, "/", "_")
  ) |>
  pivot_wider(
    names_from = student_group,
    values_from = n
  ) |>
  drop_na(hispanic_latino, white) |>
  arrange(academic_year, district, level)

new_data
```

# Part B.1: Estimating Achievement Gaps using `group_by() |> nest() |> map()`

```{r}
#| message: false

#install.packages(c("RJafroc", "yesno"))
library(RJafroc)
library(yesno)

#install.packages("gapr-master", repos = NULL, type = "source")
library(gapr)
```

```{r}
#| warning: false
v_nested_data <- new_data |>
  group_by(academic_year, district) |>
  nest() |>
  mutate(
    estimate = map(
      data,
      ~estimate_v(.x, "white", "hispanic_latino")
    )
  ) |>
  unnest(estimate) |>
  select(academic_year, district, data, auc, v, v_se) |>
  mutate(method = "group_by_nest_map")

v_nested_data
```

## Part B.1.1 Estimating achievement gaps using `nest_by() |> summarise()` 

```{r}
#| warning: false
v_rowwise <- new_data |>
  nest_by(academic_year, district) |>
  summarise(
    estimate_v(data, "white", "hispanic_latino"),
    .groups = "drop"
  ) |>
  mutate(method = "nest_by_summarise")

v_rowwise
```

# Part B.2 Confirming Same Estimates

```{r}
comparison <- v_nested_data |>
  select(academic_year, district, v_nested_data = v) |>
  left_join(
    v_rowwise |>
      select(academic_year, district, v_rowwise = v),
    by = c("academic_year", "district"))

comparison
```


```{r}
#ggplot
ggplot(comparison, aes(x = v_nested_data, y = v_rowwise)) +
  geom_point(
    color = "grey60",  
    alpha = 0.6,
    size = 2
  ) +
  geom_abline(
    intercept = 0,
    slope = 1,
    color = "#56B4E9",  
    linewidth = 1.2
  ) +
  labs(
    title = "Achievement Gap Estimates Are Equivalent Across Methods",
    subtitle = "Comparison of group_by() |> nest() |> map() and nest_by() |> summarise()",
    x = "Estimate from group_by() |> nest() |> map()",
    y = "Estimate from nest_by() |> summarise()"
  ) +
  theme_minimal()
```

# Part B.3: Creating Plots for the first 100 districts

```{r}
v_plot_data <- v_nested_data |>
  select(academic_year, district, v, v_se) |>
  left_join(
    new_data |>
      distinct(academic_year, district, n_schools),
    by = c("academic_year", "district"))

#first 100 districts
dists <- unique(v_plot_data$district)[1:100]

v_100 <- v_plot_data |>
  filter(district %in% dists)

```


```{r}
#function for one district plot
plot_gap <- function(df, district_name) {
  schools_caption <- df |>
    summarise(n_schools = max(n_schools, na.rm = TRUE)) |>
    pull(n_schools)
  
  ggplot(df, aes(x = academic_year, y = v, group = 1)) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(
      aes(ymin = v - 1.96 * v_se, ymax = v + 1.96 * v_se),
      width = 0.1
    ) +
    labs(
      title = glue("Achievement Difference: {district_name}"),
      subtitle = "Hispanic/Latino students compared with White students",
      caption = glue("Number of schools in district: {schools_caption}"),
      x = "Academic Year",
      y = "Achievement difference estimate"
    ) +
    theme_minimal()
}
```


```{r}
#Method 1: group_by() \|\> nest() \|\> map()
plots_method_1 <- v_100 |>
  group_by(district) |>
  nest() |>
  mutate(
    plot = map2(data, district, plot_gap)
  )

plots_method_1

#first plot
plots_method_1$plot[[1]]
```

## Part B.3.1 Alternative method for same plots

```{r}
#Method 2: nest_by() \|\> mutate()
plots_method_2 <- v_100 |>
  ungroup() |>
  nest_by(district) |>
  mutate(
    plot = list(plot_gap(data, district))
  ) |>
  ungroup()

plots_method_2

plots_method_2$plot[[1]]

```

# Part B.4: Save plots into `plots` directory

```{r}
#| warning: false
dir_create("plots")

# Clean file names
plots_method_1 <- plots_method_1 |>
  mutate(
    file_name = glue(
      "plots/method_1_{str_replace_all(str_to_lower(district), '[^a-z0-9]+', '_')}.png"
    )
  )

plots_method_2 <- plots_method_2 |>
  mutate(
    file_name = glue(
      "plots/method_2_{str_replace_all(str_to_lower(district), '[^a-z0-9]+', '_')}.png"))

# Save method 1 plots
walk2(
  plots_method_1$file_name,
  plots_method_1$plot,
  ~ggsave(
    filename = .x,
    plot = .y,
    width = 7,
    height = 5))

# Save method 2 plots
walk2(
  plots_method_2$file_name,
  plots_method_2$plot,
  ~ggsave(
    filename = .x,
    plot = .y,
    width = 7,
    height = 5))
```

# Extra Credit: Combine pairs of plots from both methods

```{r}
combined_plots <- plots_method_1 |>
  select(district, plot_1 = plot) |>
  left_join(
    plots_method_2 |>
      select(district, plot_2 = plot),
    by = "district"
  ) |>
  mutate(
    combined_plot = map2(
      plot_1,
      plot_2,
      ~.x / .y +
        plot_annotation(
          title = "Comparison of Plots from Two Methods",
          subtitle = "Top: group_by() |> nest() |> map(); Bottom: nest_by() |> mutate()",
          theme = theme(
            plot.title = element_text(size = 14, face = "bold"),
            plot.subtitle = element_text(size = 6)
          ))),
    file_name = glue(
      "plots/combined_{str_replace_all(str_to_lower(district), '[^a-z0-9]+', '_')}.png"))

combined_plots

# View first combined plot
combined_plots$combined_plot[[1]]


# Save combined plots
walk2(
  combined_plots$file_name,
  combined_plots$combined_plot,
  ~ggsave(
    filename = .x,
    plot = .y,
    width = 8,
    height = 9))
```
